---
title: "Basic Vol Strategy"
author: "Ben Buzzee"
date: "April 12, 2019"
output: html_document
---


This document outlines and analyzing the performance of some very basic volatility trading strategies. The idea is to use certain indicators to go short volatility.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(downloader)
require(quantmod)
require(PerformanceAnalytics)
require(TTR)
require(data.table)
library(vars)
```

# Tools of the Trade


## VIX

The VIX is the expected volatility over the next 30 days, in terms of an annualized percentage. It is derived from the prices of options for the S&P 500. To get the expected volatility for the month, multiply the VIX by sqrt(12). 

## VIX3M and VIX6M

These are like VIX except they represent the expected volatility for the next three and six months respectively, annualized.


# Ratios as Indicators

https://quantstrattrader.wordpress.com/2018/01/24/which-implied-volatility-ratio-is-best/

We will use the ratio of VIX3M/VIX6M as one of our indicators to buy. When this ratio is less than one, it means the expected volatility in the next 3 months is less than what the market expected in the next six. This ratio was chosen via backtesting.


```{r data, include = FALSE}


download("http://www.cboe.com/publish/scheduledtask/mktdata/datahouse/vix3mdailyprices.csv", 
         destfile="vxvData.csv")

# changes vxmt to vix6m to get current data
download("http://www.cboe.com/publish/scheduledtask/mktdata/datahouse/vix6mdailyprices.csv", 
         destfile="vxmtData.csv")


vxv <- xts(read.zoo("vxvData.csv", header=TRUE, sep=",", format="%m/%d/%Y", skip=2))
vxmt <- xts(read.zoo("vxmtData.csv", header=TRUE, sep=",", format="%m/%d/%Y", skip=2))


tail(vxv)


getSymbols("SPY", src = "yahoo")
getSymbols("ZIV", src = "yahoo")
getSymbols("SVXY", src = "yahoo")
getSymbols("VXX", src = "yahoo")
getSymbols("UPRO")
getSymbols("TMF")
getSymbols("XIVH")


```



```{r rets, include = FALSE}

svxyRets <- CalculateReturns(SVXY$SVXY.Adjusted)
zivRets <- CalculateReturns(ZIV$ZIV.Adjusted)
uproRets <- CalculateReturns(UPRO$UPRO.Adjusted)
tmfRets <- CalculateReturns(TMF$TMF.Adjusted)
xivhRets <- CalculateReturns(Cl(XIVH))

# use closing values
vix3mVxmt <- Cl(vxv)/Cl(vxmt)

# function to aggregate return metrics
stratStats <- function(rets) {
  stats <- rbind(table.AnnualizedReturns(rets), maxDrawdown(rets))
  stats[5,] <- stats[1,]/stats[4,]
  stats[6,] <- stats[1,]/UlcerIndex(rets)
  rownames(stats)[4] <- "Worst Drawdown"
  rownames(stats)[5] <- "Calmar Ratio"
  rownames(stats)[6] <- "Ulcer Performance Index"
  return(stats)
}

# ========= SIGNAL ================#
maLong <- SMA(vix3mVxmt, 60)


sigLong <- vix3mVxmt < 1 & vix3mVxmt < maLong

#============= RETURNS ============

# CBOE updates data at 9:45am EST on a one day lag.
# So if we check at 9:45 and the signal fires, the signal was for yesterday.
# So we can only buy at some point during the next day, and for simplicity we choose EOD.
# We only get returns 


retsShort <- zivRets
# retsVxx<-  lag(sigVxx, 1) * vxxRets
retsLongZiv <- lag(sigLong, 2) * svxyRets
retsLongSvxy <-   lag(sigLong, 2) * xivhRets
retsCombo <- lag(sigLong, 2) * (.5*svxyRets + .5*zivRets)
retsUpro <- .5*uproRets + .5*tmfRets
retsPort <- .5*retsUpro + .5*retsLongSvxy

 retsCombo <- tail(retsCombo, 300)
```



```{r}
tail(sigLong)
```


```{r perf}
compare <- na.omit(cbind(retsPort, retsLongZiv, retsLongSvxy, retsCombo))
colnames(compare) <- c("Port", "SvxyStrat", "xivHStrat", "ComboStrat")
charts.PerformanceSummary(compare)
knitr::kable(stratStats(compare))

# chart.RollingPerformance(R = retsCombo, width = 22, fun = Return.annualized)


```



```{r}
chart.CumReturns(svxyRets)
points(sigLong*25, col = "red", pch = 15)
```

```{r}
# http://volatilitymadesimple.com/chasing-the-volatility-risk-premium/



# rolling 10 day historical volatility
spyRets <- CalculateReturns(Ad(SPY))
spyVol10 <- rollapply(spyRets, width = 5, FUN = sd.annualized)
spyVol10 <- spyVol10*100


df <- merge(Cl(VIX), spyVol10, join = "inner")
colnames(df) <- c("vix", "spyvol")


rebal_bi <- seq(from = min(index(df)), to = max(index(df)), by = 30*6)
rebal_qtr <- seq(from = min(index(df)), to = max(index(df)), by = 30*3)

# Soy represented the realaized volatility of the last 10 days
# VIX is the priced-in vol for next 30

plot(df, legend.loc = "topright", grid.ticks.on = "year")

sig <- SMA(df$vix - df$spyvol, n = 10) > 0


retsSplit <- lag(sig, 2)*zivRets
retsSvxy <- lag(sig, 2)*svxyRets 
vxxRet2 <- lag(!sig, 2)*vxxRets


compare2 <- na.omit(cbind(retsSplit, retsSvxy, vxxRet2))
compare2 <- head(compare2, 800)

charts.PerformanceSummary(compare2)
stratStats(compare2)


```


```{r}
# REBALANCED VERSION

# "safe" SIV base line returns
retsLongZiv

# Dangerous SVXY returns
retsSvxy

port <- cbind(retsLongZiv, retsSvxy)
colnames(port) <- c("ziv", "svxy")
# GOAL: Juice safe returns with SVXY

rets <- Return.portfolio(port, weights = c(.7,.3), rebalance_on = "years")

stratStats(rets)
```

